---
title: "Simulation - example usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.align = 'center'
)
```




```{r setup}
library(powerHDX)
library(ggplot2)
```


Using the simulation implemented in the package *powerHDX* noisy deuteration curves can be generated. Let us consider an example peptide *PPVQ*. The function *simulate_theoretical_spectra* is used to obtain mass spectra for given exposure times provided as *times* as follows:

```{r}
times <- c(5, 10, 20, 30, 40, 50, 60, 100, 300, 500, 900, 1200, 1500, 1800)
set.seed(17)

spectra <- simulate_theoretical_spectra(sequence = "PPVQ",
                                        charge = 4,
                                        protection_factor = 100,
                                        times = times,
                                        pH = 7.5,
                                        temperature = 15,
                                        n_molecules = 500,
                                        time_step_const = 1,
                                        use_markov = TRUE)

head(spectra)
```


Simulated spectrum for particular time can be visualized using *powerHDX* as follows

```{r}
plot_single_spectrum(spectra, 30)
plot_single_spectrum(spectra, 300)
```


Deuteration curve for such a theoretical spectra is calculated via *get_deuteration_curve_single_spectrum*:

```{r}
one_curve <- get_deuteration_curve_single_spectrum(spectra)
one_curve

ggplot(one_curve, aes(x = Exposure, y = Mass)) +
  geom_point() +
  geom_line()
```


Noisy curves are generated via *get_noisy_deuteration_curves*. The number of technical replicates among which uncertainty can be calculated is *n_runs* and the number of noisy curves is *n_replicates*. 

```{r}
get_noisy_deuteration_curves(spectra, 
                             n_runs = 2,
                             n_replicates = 3,
                             compare_pairs = FALSE)
```

